---
title: 'Reproducible Research: Peer Assessment 1'
author: "Frederick Orndorff"
date: "4/26/2018"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and Preprocessing the Data

This is where we webscrape the data and store it in our local files.  
**Ensure you have the correct working directory -- the data might be saved there**

```{r}
temp <- tempfile()
git_hub_data <- "https://github.com/rdpeng/RepData_PeerAssessment1/raw/master/activity.zip"
download.file(git_hub_data, temp, mode = "wb")
unzip(temp, "activity.csv")

df <- read.csv("activity.csv")
unlink(temp) #stack overflow stated we should unlink the temp file we created

#need to change the date fields
df$date <- as.Date(df$date, format = "%Y-%m-%d")
```

## What is the Mean Total Number of Steps Taken per Day?

```{r}
total_steps_day <- aggregate(steps ~ date, df, sum)
#total_steps_day$mean_steps <- aggregate(steps ~ date, df, mean)

hist(total_steps_day[, 2],
     main = "Total Number of Steps Taken\nEach Day",
     xlab = "Total Steps per Day",
     col = "light blue",
     ylim = c(0,35))

mean(total_steps_day[, 2])
median(total_steps_day[, 2])

```

## What is the Average Daily Activity Pattern?

```{r}
avg_steps <- aggregate(steps ~ interval, df, mean)
plot(x = avg_steps[, 1],
     y = avg_steps[, 2],
     type = "l",
     main = "Average Number of Steps\nper 5-minute interval",
     xlab = "5-minute interval",
     ylab = "Average Number of Steps",
     col = "blue",
     lwd = 2.5)
avg_steps[which(avg_steps[, 2] == max(avg_steps[, 2])), 1]
```

## Imputing Missing Values

```{r}
sum(is.na(df$steps))

library("stats")

df$steps_new <- ave(df$steps, df$interval,
                    FUN = function(x) replace(x, is.na(x), mean(x, na.rm = TRUE)))

df_new <- df[, c(4, 2:3)]

total_steps_day2 <- aggregate(steps_new ~ date, df_new, sum)

hist(total_steps_day2[, 2],
     main = "Total Number of Steps Taken\nEach Day",
     xlab = "Total Steps per Day",
     col = "light blue")

mean(total_steps_day2[, 2])
median(total_steps_day2[, 2])
```

## Are there Differences in Activity Patterns Between Weekdays and Weekends?

```{r}
weekdays_factor <- c("Monday", "Tuesday", "Wednesday",
                     "Thursday", "Friday")
df_new$wk_day <- factor((weekdays(df_new$date) %in% weekdays_factor),
                        levels = c(FALSE, TRUE),
                        labels = c("weekend", "weekday"))


df_wk_day <- df_new[df_new$wk_day == "weekday", ]
df_wkend_day <- df_new[df_new$wk_day == "weekend", ]

avg_steps_wkday <- aggregate(steps_new ~ interval, df_wk_day, mean)
avg_steps_wkend <- aggregate(steps_new ~ interval, df_wkend_day, mean)

par(mfrow = c(2,1))

plot(x = avg_steps_wkday[, 1],
     y = avg_steps_wkday[, 2],
     type = "l",
     main = "Average Number of Steps per 5-minute interval\nWeekday",
     xlab = "5-minute interval",
     ylab = "Average Number of Steps",
     col = "green",
     lwd = 2.5)

plot(x = avg_steps_wkend[, 1],
     y = avg_steps_wkend[, 2],
     type = "l",
     main = "Average Number of Steps per 5-minute interval\nWeekend",
     xlab = "5-minute interval",
     ylab = "Average Number of Steps",
     col = "light green",
     lwd = 2.5)
```
